BlueTracer: \\ a Robust API Tracer for Evasive Malware

### 2018-06-23 LMThesisNicchi.tex
Old state of the art stuff removed

\chapter{BlueTracer}
So far only userland techniques.
-> There are also kernel hooks ...
\\
3 categories of hooks: binary rewriting, VM based and DBI based.

\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}

	\hspace*{-3.3cm}
    \begin{tabular}{ | M{2cm} | M{4cm} | M{4cm} | M{4cm} | M{3cm}|}
    \hline
    \textbf{API Hooking Technique} & \textbf{Brief Overview} & \textbf{Strengths} & \textbf{Weaknesses} & \textbf{Analysis tools which use it} \\ \hline
    \textbf{Inline hooking} & The API to be hooked has its initial instructions overwritten with an unconditional jump to a replacement function. In order to ensure that the API's original functionality is not lost due to the modification of its entry point, a \textit{trampoline function} is created, consisting of a copy of the overwritten instructions and an unconditional jump back to the unaltered portion of the original function. As a result of this, the replacement function can invoke the original function by calling the trampoline. 
\iffalse
The required program modifications are carried out either before the binary is loaded and executed, through the addition of supplementary sections, or while the binary is already running via DLL injection.
\fi    & 
    \begin{itemize}
	\item Fast and efficient.
    \item Can be utilized to hook any code, not just operating systems API functions but also programmer defined functions. Furthermore, unlike IAT patching, the type of command used to call the function does not matter, meaning that the hooking will be effective regardeless of the fact that a function is called using the IAT or using \texttt{LoadLibrary} together with \texttt{GetProcAddress}.  
    \end{itemize}.  
    & \begin{itemize}
	\item Needs additional modifications in the case where the function's entry point includes specific instructions, like ones which contain relative memory addresses. In fact, such instructions cannot be executed from a trampoline as the trampoline is located in a different memory location than the one of the original program code.
    \item Can be easily detected, for instance by comparing the code section of system libraries in memory with a matching original copy loaded from the file system to detect library modifications or by searching API entry points for specific patterns (e.g. PUSH-RET combo).
    \end{itemize}  
    & \begin{itemize}
	\item Microsoft Detours
	\item EasyHook 
    \item Cuckoo Sandbox
    \item WinAPI Override32
    \end{itemize}  
     \\ \hline
\end{tabular}

	\hspace*{-3.3cm}
    \begin{tabular}{ | M{2cm} | M{4cm} | M{4cm} | M{4cm} | M{3cm}|}
    \hline
    \textbf{API Hooking Technique} & \textbf{Brief Overview} & \textbf{Strengths} & \textbf{Weaknesses} & \textbf{Analysis tools which use it}
     \\ \hline
    \textbf{Import Address Table (IAT) Patching} & In the header of every Portable Executable (PE) file there is an IAT for every DLL that is included by the executable. This table is utilized to indicate the location of DLL-imported functions in virtual memory and is filled by the Windows loader with the actual function memory addresses after the executable is loaded in memory. The idea is to overwrite the original pointer to an imported API function so that, instead of pointing to the original API, it will point to a different function. & 
    \begin{itemize}
    \item Simple and easy to implement
	\end{itemize}     
    & 
    \begin{itemize}
    \item Old and extremely easy to detect by simply examining the entries of the IAT and checking whether or not each address falls inside the memory range of the DLL that should contain the function.
    \item It is necessary to parse every IAT of every imported DLL to verify that the target API is actually imported.
    \item Ineffective when function pointers are acquired dynamically, e.g. via \texttt{LoadLibrary} and \texttt{GetProcAddress}.
    \end{itemize}
    & None    
      \\ \hline
    \textbf{Export Address Table (EAT) Patching} & Similar to IAT patching, but, this time, DLL export address tables are patched instead. The export address table (EAT) contains the name of every function exported by the DLL together with the relative virtual address (RVA) where the function can be found, which is relative to the DLL base address when loaded in memory. To hook an API function via EAT patching all that is needed is to overwrite the corresponding address in the table with the address of another function.   &
    \begin{itemize}
    \item Produces similar results to the ones obtained through IAT patching, but, unlike IAT patching, the created hooks are global, i.e. they affect every program which utilizes the altered DLL.
	\end{itemize} 
     & 
    \begin{itemize}
    \item Similarly to IAT patching, it can be easily detected to by simply examining the entries of the EAT and checking whether or not each RVA, when added to the DLL base address, falls within the DLL memory range.
    \end{itemize}
	 & Joebox \\
    \hline
    \end{tabular}

	\hspace*{-3.3cm}
    \begin{tabular}{ | M{2cm} | M{4cm} | M{4cm} | M{4cm} | M{3cm}|}
    \hline
    \textbf{API Hooking Technique} & \textbf{Brief Overview} & \textbf{Strengths} & \textbf{Weaknesses} & \textbf{Analysis tools which use it}
     \\ \hline
    \textbf{Proxy DLL (or Trojan DLL)} & The DLL containing the functions to be hooked is replaced with another one having an identical name and exporting all the symbols of the original DLL. In addition to calling the original functions so that they can carry out their tasks, the proxy DLL may also make available different implementations for the hooked functions.    & 
    \begin{itemize}
    \item Trivial to implement
	\end{itemize}     
    & 
    \begin{itemize}
    \item Extremely easy to detect since the original DLL is substituted with another file, which is very likely to have a different size. Moreover, checksums could be employed to detect the presence of a trojan DLL.  
    \end{itemize}
    & Norman Sandbox    
      \\ \hline
    \textbf{Using a debugger} & Hooking through the use of a debugger is realized by instructing the debugger to position a breakpoint at the entry point of the target API function. The placement of a breakpoint involves overwriting the initial instructions of the target API functions with CPU specific instructions, like INT 3 for IA-32. These lead the CPU to throw a debug exception in case they are pointed by the current instruction pointer (IP). The exception is then intercepted by the debugger, which is able to deduce the API which is being called by the application from the address at which the exception took place. Moreover, the debugger also has total control over the memory contents and the CPU state of the process being debugged.     &
    \begin{itemize}
    \item Contrarily to inline hooking, it can be used to hook functions whose entry points include instructions containing relative addresses.
	\end{itemize} 
     & 
    \begin{itemize}
    \item Debug exceptions suspend all the threads of the application.
    \item There has to be a debugger process handling the exception and the target program is typically loaded as a child process inside the debugger. A malware can check for this at run-time.
    \item Bad performance
    \item Extremely easy to detect as the existence of the INT 3 instruction in an API entry point immediately gives away the debugger's presence
    \item Many malwares employ techniques to check for the presence of a debugger.
    \end{itemize}
	 & Any debugger, e.g. WinDbg \\
    \hline
    \end{tabular}

	\hspace*{-3.3cm}
    \begin{tabular}{ | M{2cm} | M{4cm} | M{4cm} | M{4cm} | M{3cm}|}
    \hline
    \textbf{API Hooking Technique} & \textbf{Brief Overview} & \textbf{Strengths} & \textbf{Weaknesses} & \textbf{Analysis tools which use it} \\ \hline
    \textbf{VMI-based hooking} & The target program is executed in an emulated environment, typically with QEMU being used as virtual machine monitor (VMM). Function calls are monitored by comparing the virtual processor's instruction pointer with the RVAs of DLLs' exported functions when added to the DLL base address. Function arguments are also monitored and this is done by providing them to callback routines, which perform the appropriate tracking operations.       & 
    \begin{itemize}
	\item A PC emulator allows to have functionalities similar to the ones of a debugger, i.e. the code being monitored can be stopped at any arbitrary point during its execution, allowing its registers and virtual memory to be inspected. However, unlike a debugger, it is not subject to the aforementioned issues related to breakpoints.
    \item Harder to detect since emulation is utilized to execute an unknown binary with a complete operating system in software, without the sample being never ran directly on the processor.
    \end{itemize}.  
    & \begin{itemize}
	\item Incurs in the \textit{semantic gap} problem, i.e. the issue of deducing high-level information from the raw system information by making sense of the CPU state and memory contents.
	\item Although it requires more work, it is still possible for a sample to discover side effects of the emulation.     
    \end{itemize}  
    & \begin{itemize}
	\item TTAnalyze
	\item TQana
	\item Pyrebox
    \end{itemize}  
     \\ \hline
\end{tabular}

>>>>

\chapter{BlueTracer}
